---
title: "EU Population over 65"
author: "Raquel Baltazar"
date: "`r format(Sys.time(), '%a %d %b %Y (%H:%M:%S)')`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

**We will use Eurostat’s indicator data for “Proportion of population aged 65 and over” as a case study when working with the data.table R package and apply a few animations using gganimate. The goal is to explore the geographical and time trends for the gender pay gap in the EU and compare Portugal with its countries**


We will start by learning how to understand the data and extract it from the Eurostat package within R. We will then process in preparation for some exploratory analysis.*

# Objective
The objective is to look at the geographical and time trends in the data. We will
answer the following questions:

+ What are the time trends for Portugal?
+ How does Portugal compare to other European countries?
+ Which countries have the highest proportion of the population being 65 and over in Europe by year?

# Understanding the Data

##  Proportion of population aged 65 and over
% of total population who are aged 65 and over 

## Source 
online data code: TPS00028
Source of data: Eurostat
Last data update: 03/07/2020  22:00 (5 months ago)
Last structure update: 03/07/2020
Overall data coverage: 2008 — 2019
Number of values: 592

source link: https://ec.europa.eu/eurostat/databrowser/view/tps00028/default/table?lang=en
The source dataset for TPS00028 is:
DEMO_PJANIND
https://ec.europa.eu/eurostat/databrowser/product/view/DEMO_PJANIND

**Copyrights:** Eurostat Copyright/Licence Policy is applicable.


# Loading Libraries

```{r library}
library(eurostat)
library(knitr)
library(kableExtra)
library(ggplot2)
library(magrittr)
library(data.table)
library(here)
library(countrycode)
library(gganimate)
library(gapminder)
library(data.table)
library(dplyr)
```

# Importing the Data

```{r query}

query <- search_eurostat(pattern = "Proportion of population aged 65 and over", 
                         type = "table", fixed = FALSE)
query[, 1:2]
```

```{r ct_and_ptandeu}

ct <-  c("AT", "BE", "BG", "CH", "CY", "CZ", "DE", "DK", "EE", "EL", "ES", "FI", "FR", "HR","HU", "IE", "IS", "IT", "LI", "LT","LU", "LV", "MT", "NL", "NO", "PL", "PT","RO", "SE", "SI", "SK", "UK")

PTandEU <-  c("PT","EU27_2020")
```

# Defining variables dat and dat1

## dat = Every country in the EU separately
## dat1 = Portugal vs EU

```{r get_eurostat}


dat <- get_eurostat(id="tps00028", time_format = "num", 
                    filters = list (geo = ct))
dat [1:2,]

dat1<- get_eurostat(id="tps00028", time_format = "num", 
                    filters = list (geo = PTandEU))
                    
dat1 [1:2,]

```


```{r label}
dat <- label_eurostat (dat)
dat[1:3,]


dat1 <- label_eurostat (dat1)
dat1[1:3,]
```


# Evolution of proportion of the population 65 and over, from 2008 to 2019

```{r eurostat_and_plots}

ggplot(dat, 
       aes(x = time, y= values, color = geo, label = geo)) + 
  geom_line (alpha = .5) +
  geom_text (data = dat %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  labs(title = "Proportion of people 65 and over",
       x= "Year", y= "%")
```

## ...The same line graph, but animated

```{r animated_line}

ggplot(dat, aes(time, values, color = geo, label = geo)) +
  geom_line(alpha = .5) +
  geom_text (data = dat %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  scale_color_viridis_d() +
  labs(x = "Year", y = "Proportion of people 65 and over") +
  theme(legend.position = "top") + 
  labs(title = "Proportion of people 65 and over",
       x= "Year", y= "%") +
  transition_reveal(time) +
  geom_point() +
  transition_reveal(time)
```

## ... in an animated bar chart 

```{r animated_bar_chart1}

ggplot (dat, aes(x= reorder(geo, values), y = values)) + 
  geom_col (color = "white", fill = "grey80") + 
  theme ( axis.text.x = element_text (size = 6)) + 
  labs( title = "Proportion of people 65 and over",
        subtitle = "Percentage of the total population",
        fill = "%") + 
  labs(title = 'Year: {frame_time}') +
  transition_time(time) +
  ease_aes()
```

# Portugal vs. European Union, 2008-2019

## in a line graph
```{r PT_vs_EU}

pteu <-  c("European Union - 27 countries (from 2020)","Portugal")

library(ggplot2)
library(dplyr)
ggplot(dat1, 
       aes(x = time, y= values, color = geo, label = geo)) + 
  geom_line (alpha = .5) +
  geom_text (data = dat1 %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  labs(title = "Proportion of people 65 and over",
       x= "Year", y= "%")
```

## ... The same line graph, but animated

```{r ptvseu_animated}

ggplot(dat1, aes(time, values, color = geo, label = geo)) +
  geom_line(alpha = .5) +
  geom_text (data = dat1 %>% group_by(geo) %>% 
               filter(time == max(time)),
             size =2.6) +
  theme(legend.position = "none") + 
  scale_color_viridis_d() +
  labs(x = "Year", y = "Gender Pay Gap") +
  theme(legend.position = "top") + 
  labs(title = "Proportion of people 65 and over",
       x= "Year", y= "%") +
  transition_reveal(time) +
  geom_point() +
  transition_reveal(time)


```

# Then vs Now (bar chart and map)
Now we compare the first and the latest years [2008 vs. 2019] provided in the dataset individually.  

## Proportion of people 65 and over in 2008

### ...in a Bar Chart (2008)

```{r bar_chart}
dat_2008 <- dat %>% 
  filter(time == "2008")
ggplot (dat_2008, aes(x= reorder(geo, values), y = values)) + 
  geom_col (color = "white", fill = "grey80") + 
  theme ( axis.text.x = element_text (size = 6)) + 
  labs (title = "Proportion of people 65 and over in 2008", 
        y = "%", x = NULL)
```

### ...in a Map (2008)
```{r map}
mapdata <-  get_eurostat_geospatial(nuts_level = 0) %>% 
  right_join (dat_2008) %>% 
  mutate (cat = cut_to_classes (values, n = 4, decimals = 1))
head(select(mapdata,geo,values,cat), 3)

```

```{r plot_map}

ggplot(mapdata, aes(fill = cat)) + 
  scale_fill_brewer(palette = "RdYlBu") + 
  geom_sf (color = alpha("white", 1/3), alpha = .6) + 
  xlim (c(-12,44)) + ylim(c(35, 70)) + 
  labs( title = "Proportion of people 65 and over in 2008",
        fill = "%")
```

# Proportion of people 65 and over in 2019

## ...in a bar chart
```{r bar_chart_19}

dat_2019 <- dat %>% 
  filter(time == "2019-01-01")
ggplot (dat_2019, aes(x= reorder(geo, values), y = values)) + 
  geom_col (color = "white", fill = "grey80") + 
  theme ( axis.text.x = element_text (size = 6)) + 
  labs (title = "Proportion of people 65 and over in 2019", 
        y = "%", x = NULL)
```

# ...in a map
```{r map_19}

mapdata <-  get_eurostat_geospatial(nuts_level = 0) %>% 
  right_join (dat_2019) %>% 
  mutate (cat = cut_to_classes (values, n = 4, decimals = 1))
head(select(mapdata,geo,values,cat), 3)

```


```{r plot_map_19}
ggplot(mapdata, aes(fill = cat)) + 
  scale_fill_brewer(palette = "RdYlBu") + 
  geom_sf (color = alpha("white", 1/3), alpha = .6) + 
  xlim (c(-12,44)) + ylim(c(35, 70)) + 
  labs( title = "Population above 65 in 2019",
        subtitle = "Percentage of the population aged 65 or more in 2019",
        fill = "%")
```


